---
title: "My Awesome Report"
author: "Péter Király"
date: "2024-11-27"
output: html_document
---
## Introduction
bold: **bold** or __bold__

italics: *italics*, or _italics_

bold and italics: ***really***, or ___really___ or **_really_**, **_really_**.

code-type: `code-type`

a list:

- village
- interview_date
- no_members
- years_liv
- respondent_wall_type
- rooms

ordered list:

1. village
2. interview_date
3. no_members
4. years_liv
5. respondent_wall_type
6. rooms

And nested items by tab-indenting:

- village
  + Name of village
- interview_date
  + Date of interview
- no_members
  + How many family members lived in a house
- years_liv
  + How many years respondent has lived in village or neighbouring village
- respondent_wall_type
  + Type of wall of house
- rooms
  + Number of rooms in house

Reference: https://www.markdownguide.org/basic-syntax

R code:

```{r chunk-name}
"Here is where you place the R code that you want to run."
```

a chunk called "setup". Do not show the output

```{r setup, include = FALSE}
library(tidyverse)
library(here)
interviews <- read_csv(here("data/SAFI_clean.csv"), na = "NULL")
```

# Insert table

```{r table, echo = FALSE}
interviews %>%
    filter(!is.na(memb_assoc)) %>%
    group_by(village, memb_assoc) %>%
    summarize(mean_no_membrs = mean(no_membrs)) %>%
  knitr::kable(caption = "We can also add a caption.", 
               col.names = c("Village", "Member Association", 
                             "Mean Number of Members"))
```

# Customising chunk output

- eval (TRUE or FALSE): Whether or not the code within the code chunk should be run.
- echo (TRUE or FALSE): Choose if you want to show your code chunk in the output document. echo = TRUE will show the code chunk.
- include (TRUE or FALSE): Choose if the output of a code chunk should be included in the document. FALSE means that your code will run, but will not show up in the document.
- warning (TRUE or FALSE): Whether or not you want your output document to display potential warning messages produced by your code.
- message (TRUE or FALSE): Whether or not you want your output document to display potential messages produced by your code.
- fig.align (default, left, right, center): Where the figure from your R code chunk should be output on the page


Exercise

Play around with the different options in the chunk with the code for the table, and re-Knit to see what each option does to the output.

What happens if you use eval = FALSE and echo = FALSE? What is the difference between this and include = FALSE?

# In-line R code

today’s date is `r Sys.Date()`, 

```{r}
# create a summary data frame with the mean household size by village
mean_household <- interviews %>%
    group_by(village) %>%
    summarize(mean_no_membrs = mean(no_membrs))

# and select the village we want to use
mean_chirodzo <- mean_household %>%
  filter(village == "Chirodzo")
```

The average household size in the village of Chirodzo is `r round(mean_chirodzo$mean_no_membrs, 2)`

# Plots

```{r}
## Not run, but can be used to load in data from previous lesson!
interviews_plotting <- interviews %>%
  ## pivot wider by items_owned
  separate_rows(items_owned, sep = ";") %>%
  ## if there were no items listed, changing NA to no_listed_items
  replace_na(list(items_owned = "no_listed_items")) %>%
  mutate(items_owned_logical = TRUE) %>%
  pivot_wider(names_from = items_owned, 
              values_from = items_owned_logical, 
              values_fill = list(items_owned_logical = FALSE)) %>%
  ## pivot wider by months_lack_food
  separate_rows(months_lack_food, sep = ";") %>%
  mutate(months_lack_food_logical = TRUE) %>%
  pivot_wider(names_from = months_lack_food, 
              values_from = months_lack_food_logical, 
              values_fill = list(months_lack_food_logical = FALSE)) %>%
  ## add some summary columns
  mutate(number_months_lack_food = rowSums(select(., Jan:May))) %>%
  mutate(number_items = rowSums(select(., bicycle:car)))
```


Exercise

Create a new code chunk for the plot, and copy the code from any of the plots we created in the previous episode to produce a plot in the chunk. I recommend one of the colourful plots.

If you are feeling adventurous, you can also create a new plot with the interviews_plotting data frame.

create a caption with the chunk option fig.cap.

```{r plot, fig.cap = "I made this plot while attending an awesome Data Carpentries workshop where I learned a ton of cool stuff!"}
interviews_plotting %>%
  ggplot(aes(x = respondent_wall_type)) +
  geom_bar(aes(fill = village), position = "dodge") + 
  labs(x = "Type of Wall in Home", y = "Count", fill = "Village Name") +
  scale_fill_viridis_d() # add colour deficient friendly palette
```